Fractional anisotropy as an optimal tracer of cosmic voids

### 2014-08-10 cosmic_voids.tex
Finishing introduction

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Fractional anisotropy as an optimal tracer of cosmic voids}
\author[S. Bustamante and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Sebastian Bustamante \thanks{sbustamante@udea.edu.co}$^{1}$,
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^1$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia\\ 
$^2$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

\begin{abstract}
In the last decades voids have been identified as one of the most 
important features of the Megaparsec Universe, therefore a full 
understanding of their properties is required in order to a proper 
description of the large-scale structure of the Cosmic Web. Identifying
voids in cosmological observations and N-body simulations is hence an 
important enterprise that must be carried out and accordingly, a plethora 
of different void finding techniques can be found in the literature.

@@

We test our method on the Bolshoi simulation and compare it with a
density-based scheme. We find our method gives a better resemblance to the
visual impression of the Cosmic Web in addition to a more robust 
description of voids as it allows to identify, in a very clear manner, 
their spatial limits as regions where the FA is maximized, normally 
corresponding to walls and filaments. This is mainly because of the 
non-monotonic behaviour of the FA, what allows to use it as a sort of 
one-dimensional tomography of the Cosmic Web.

@@

Since voids where discovered in the first compiled galaxy surveys 
\citep{Chincarini75, Gregory78, Einasto80M, Einasto80N, Kirshner81, 
Kirshner87}, they have been identified, along with the filamentary and 
hierarchically clustered nature of the Cosmic Web, as one of the most 
striking features of the Megaparsec Universe \citep{Bond96}. Nevertheless,
due to the large volume extension of void regions ($\sim 5-10\ \mbox{Mpc} 
h^{-1}$), statistically significant catalogues of voids \citep{Pan10, 
Sutter12b, Nadathur14} only have become available after modern large 
galaxy surveys like the two-degree field Galaxy Survey \citep{Colless01,
Colless03} and the Sloan Digital Sky Survey \citep{York00,Abazajian03}, 
thereby triggering a plethora of more refined observational and 
statistical studies of voids throughout the last decade \citep{Hoyle04,
Croton04, Rojas05, Ceccarelli06, Patiri06a, Tikhonov06, Patiri06b, 
Tikhonov07, BendaBeckmann08, Foster09, Ceccarelli13, Sutter14a}

On the theoretical side, early descriptions of the evolution of the 
large-scale Universe, based on gravitational instabilities in primordial 
stages and leaded by the seminal work of \citet{Zeldovich70}, are highly 
consistent with the Cosmic Web picture, where planar pancake-like regions 
of matter enclose enormous sub-dense voids and are bordered, in turn, by 
thin filaments and high-density clumpy knots \citep{Bond96}. First 
theoretical models for describing formation, dynamics and properties of 
voids \citep{Hoffman82, Icke84, Bertschinger85, Blumenthal92} were 
dramatically complemented and extended by first numerical studies based 
on simulations \citep{Martel90, Regos91, Weygaert93, Dubinski93}. This 
tendency of using numerical data from N-body simulations, fuelled by last 
generation computing systems and ever more efficient numerical algorithms, 
has become increasingly common in the last years as a powerful analysis 
toolkit of cosmic voids (for an extensive compilation of previous 
numerical works, see \citet{Colberg08}).

At present, studying voids may be considered as a threefold enterprise
\citep{Platen07}: first, they are a key ingredient of the Cosmic Web as 
they dominate almost the entire volume distribution at large-scales and 
additionally, they compensate over-dense structures in the total budget of 
matter. This implies that a full understanding of their evolution and 
properties is essential for fathoming the high complexity of the Cosmic 
Web. Second, they provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. Third, they constitute and unique and still 
largely pristine environment in which can be tested and probed theories 
of galaxy formation and evolution.

Although a visual recognition of voids on galaxy surveys and simulations
is possible in most cases, a formal systematic identification of them 
is necessary for statistically meaningful studies. However, a basic yet 
essential issue remains regarding the definition itself of what a void is. 
There is not a general consensus about an unambiguous definition of cosmic 
voids and therefore, there can be found many different void finding 
techniques throughout the literature (for a detailed comparison of 
different schemes, see the Void Finder Comparison Project 
\citet{Colberg08}). In spite of the diversity of existing schemes, they 
can be roughly classified into two main types: first, schemes based on 
spatial or redshift distribution of galaxies in surveys or catalogues of 
dark matter halos in simulations \citep{Kauffmann91, Muller00, Gottlober03,
Hoyle04, Brunino07, Foster09, Micheletti14, Sutter14}, and second, schemes 
based on the smooth and continuous matter density field either from 
simulations or reconstruction procedures on surveys \citep{Plionis02, 
Colberg05, Shandarin06, Platen07, Neyrinck08, Neyrinck13, Ricciardelli2013}

@@

For the distribution of the density with respect to the $\lambda_1$ 
eigenvalue, it can be appreciated an analogous behaviour, where central 
regions of voids present the most under-dense values of the overall cosmic 
web. For outer layers of voids, the density field grows progressively 
until sheets and filaments are reached. However, the dispersion in the 
distributions indicates that the geometry of voids as quantified by both 
web schemes is not compatible with the density field, i.e. contours of the
$\lambda_1$ eigenvalue do not coincide with contours of the density field. 
Then, the density is not a reliable quantity to be used as tracer of our 
voids. Furthermore, another advantage of using the FA index instead of the
density is the non-monotonous behaviour of the FA median value, with a 
local maxima that allows to identify properly the boundaries of voids.

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Fractional anisotropy as an optimal tracer of cosmic voids}
\author[S. Bustamante and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Sebastian Bustamante \thanks{sebastian.bustamante@udea.edu.co}$^{1}$,
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^1$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia\\ 
$^2$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

\begin{abstract}
In the last decades voids have been identified as one of the most 
important features of the Megaparsec Universe, therefore a full 
understanding of their properties is required in order to a proper 
description of the large-scale structure of the Cosmic Web. Identifying
voids in cosmological observations (e.g. redshift galaxy surveys) and 
N-body simulations is hence an important enterprise that must be carried 
out and accordingly, a plethora of different void finding techniques can 
be found in the literature.

@@

We test our method on the Bolshoi simulation and compare with a
density-based scheme. We find our method gives a better resemblance of 
the visual impression of the distribution of voids within the Cosmic Web, 
in addition to a more robust description of voids as it allows to identify,
in a very clear manner, their outlines as regions where the FA is 
maximized, normally corresponding to walls and filaments. This is mainly 
because of the non-monotonic behaviour of the FA, what allows to use it 
as a sort of one-dimensional tomography of the Cosmic Web.

@@

Since voids where discovered in the first compiled galaxy surveys 
\citep{Chincarini75, Gregory78, Einasto80M, Einasto80N, Kirshner81, 
Kirshner87}, they have been identified, along with the filamentary and 
hierarchically clustered nature of the Cosmic Web, as one of the most 
striking features of the Megaparsec Universe \citep{Bond96}. Nevertheless,
due to the large volume extension of void regions ($\sim 5-10\ \mbox{Mpc} 
h^{-1}$), statistically significant catalogues of voids \citep{Pan10, 
Sutter12b, Nadathur14} only have become available after modern large 
galaxy surveys like the two-degree field Galaxy Redshift Survey \citep{
Colless01, Colless03} and the Sloan Digital Sky Survey \citep{York00,
Abazajian03}, thereby triggering a plethora of more refined observational 
and statistical studies of voids throughout the last decade \citep{Hoyle04,
Croton04, Rojas05, Ceccarelli06, Patiri06a, Tikhonov06, Patiri06b, 
Tikhonov07, BendaBeckmann08, Foster09, Ceccarelli13, Sutter14a}

On the theoretical side, early descriptions of the evolution of the 
large-scale Universe, based on gravitational instabilities in primordial 
stages and leaded by the seminal work of \citet{Zeldovich70}, are highly 
consistent with the Cosmic Web picture, where planar pancake-like regions 
of matter enclose enormous underdense voids and are bordered, in turn, by 
thin filaments and high-density clumpy knots \citep{Bond96}. First 
theoretical models for describing formation, dynamics and properties of 
voids \citep{Hoffman82, Icke84, Bertschinger85, Blumenthal92} were 
dramatically complemented and extended by first numerical studies based 
on simulations \citep{Martel90, Regos91, Weygaert93, Dubinski93}. This 
tendency of using numerical data from N-body simulations, fuelled by last 
generation computing systems and ever more efficient numerical algorithms, 
has become increasingly common in the last years as a powerful analysis 
toolkit of cosmic voids (for an extensive compilation of previous 
numerical works, see \citet{Colberg08}).

At present, studying voids may be considered as a threefold enterprise
\citep{Platen07}: first, they are a key ingredient of the Cosmic Web as 
they dominate almost the entire volume distribution at large-scales and 
additionally, they compensate overdense structures in the total budget of 
matter. This implies that a full understanding of their evolution and 
properties is essential for fathoming the high complexity of the Cosmic 
Web. Second, they provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. Third, they constitute and unique and still 
largely pristine environment in which can be tested and probed theories 
of galaxy formation and evolution.

Although a visual recognition of voids on galaxy surveys and simulations
is possible in most cases, a formal systematic identification of them 
is necessary for statistically meaningful studies. However, a basic yet 
essential issue remains regarding the definition itself of what a void is. 
There is not a general consensus about an unambiguous definition of cosmic 
voids and therefore, there can be found many different void finding 
techniques throughout the literature (for a detailed comparison of 
different schemes, see the Void Finder Comparison Project 
\citet{Colberg08}). In spite of the diversity of existing schemes, they 
can be roughly classified into two main types: first, geometric schemes 
based on point spatial or redshift distribution of galaxies in surveys or 
catalogues of dark matter halos in simulations \citep{Kauffmann91, 
Muller00, Gottlober03, Hoyle04, Brunino07, Foster09, Micheletti14, 
Sutter14}, and second, schemes based on the smooth and continuous matter 
density field either from simulations or from reconstruction procedures 
on surveys \citep{Plionis02, Colberg05, Shandarin06, Platen07, Neyrinck08,
Neyrinck13, Ricciardelli2013}.

Here we introduce a new tracer of the structure of cosmic voids built from 
two tensorial web schemes, i.e. the T-web, based on the Hessian of the 
gravitational potential or tidal tensor \citep{Hahn07, Forero09}, and the 
V-web, based on the velocity shear tensor \citep{Hoffman12}. These web 
schemes classify the Cosmic Web into four different types of environment
depending on the counting of the number of eigenvalues below an 
user-defined threshold ($\lambda_{th}$), namely voids (3 eigenvalues below
$\lambda_{th}$), sheets (2), filaments (1) and knots or clusters (0).
The tidal and the shear tensors have proven to be more fundamental than 
the density field as they also trace the collapsing or expanding nature of 
the matter field, which ultimately is what defines the underlying dynamics 
of the Cosmic Web. Furthermore, the density field is degenerated regarding 
the defined types of environment \citep{Hahn07}, which implies that is not 
possible to assign a priori a range of density values to each of them. 
This yields that the usually adopted definition of voids as underdense 
regions in the large-scale matter distribution, is not precise enough as 
it excludes a proper description of the internal structure of voids.

Following the recent trend of using digital image-processing techniques 
developed in other disciplines (e.g. medical sciences and computer imaging) 
for analysing the structure of the Cosmic Web, we propose here, much in 
the same way as \citet{Libeskind13}, the fractional anisotropy (FA) as a 
tracer of the internal structure and the outline of cosmic voids. The FA 
was initially introduced by \citet{Basser95} for quantifying the 
anisotropy degree of the diffusivity of water molecules through cerebral
tissue in nuclear magnetic resonance imaging. In this context we use the 
same quantity for determining the anisotropy degree of the local 
environment from either the orbital dynamics as set by the eigenvalues of
the tidal tensor (T-web), or the dynamics of the velocity field as set by 
the eigenvalues of the shear tensor (V-web). In both cases, the FA is not
determined directly from the density field, hence it is suitable for 
describing the structure of high non-linear regions (e.g. filaments, 
knots and very dense walls) as well as for revealing the fainter 
substructure exhibited by quasi-linear regions like voids.

Once established the FA as the tracer field of cosmic voids, we proceed to
identify individual voids as basins of local minima. For this purpose we
implement a technique defined by \citet{Beucher79} and \citet{Beucher93} 
and widely used in the field of image analysis and mathematical morphology, 
i.e. the \textit{watershed transform algorithm}. This technique has been 
already implemented on void finding schemes by \citet{Platen07} and 
\citet{Neyrinck08} with very interesting results, where individual voids 
are identified as catching basins of local minima of the density field. 
The attractiveness of this algorithm lies in it is parameter-free and does 
not require to make any assumption on the shape and morphology of voids. 
Although we just use a \textit{cloud-in-cell} (\texttt{CIC}) algorithm on 
a Cartesian mesh for estimating the density and tensor fields, instead of 
the more sophisticated \textit{Delaunay tessellation for field estimator} 
(\texttt{DTFE}) technique \citep{Schaap00}, our implementation of the 
watershed transform should not be significantly affected as we are 
interested in quasi-linear regions where the \texttt{CIC} gives similar 
estimations.

This paper is organized as follows.

@@

For the distribution of the density with respect to the $\lambda_1$ 
eigenvalue, it can be appreciated an analogous behaviour, where central 
regions of voids present the most underdense values of the overall cosmic 
web. For outer layers of voids, the density field grows progressively 
until sheets and filaments are reached. However, the dispersion in the 
distributions indicates that the geometry of voids as quantified by both 
web schemes is not compatible with the density field, i.e. contours of the
$\lambda_1$ eigenvalue do not coincide with contours of the density field. 
Then, the density is not a reliable quantity to be used as tracer of our 
voids. Furthermore, another advantage of using the FA index instead of the
density is the non-monotonous behaviour of the FA median value, with a 
local maxima that allows to identify properly the boundaries of voids.
